Collaboration in Ad Hoc Teamwork: \\ Ambiguous Tasks, Roles, and Communication

### 2016-01-31 paper/adhoccom.tex
formatting for http://ala2016.csc.liv.ac.uk/

\documentclass[a4paper]{article}

\usepackage{conference}
\usepackage{times}

@@

\title{Collaboration in Ad Hoc Teamwork: \\
Ambiguous Tasks, Roles, and Communication\vspace{1em}}

% \author{Anonymous authors for blind review}
\setlength\titlebox{3.8in}

\author{
Jonathan Grizou \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ jonathan.grizou@inria.fr
\And
Samuel Barrett \\ Kiva Systems \\ North Reading, MA 01864 USA \\ basamuel@kivasystems.com
\AND
Peter Stone \\ Dept. of Computer Science \\ The Univ. of Texas at Austin \\ Austin, TX 78712 USA \\ pstone@cs.utexas.edu
\And
Manuel Lopes \\ Flowers Team \\ INRIA - ENSTA ParisTech \\  France \\ manuel.lopes@inria.fr
}

\newcommand{\env}{E}
\newcommand{\task}{\tau}
\newcommand{\conf}{\kappa}
\newcommand{\com}{\rho}
\newcommand{\gridwidth}{1}
\newcommand{\figwidth}{1}

@@

\maketitle

\begin{abstract}
% Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents.
Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. We consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that our ad hoc agent can join such a team without barely impacting the overall performance compared to a pre-coordinated agent.
\end{abstract}

@@

\bibliographystyle{named}
\bibliography{adhoccom}

>>>>

\documentclass{aamas2016}

\pdfpagewidth=8.5truein
\pdfpageheight=11truein

\usepackage{times}

@@

\title{Collaboration in Ad Hoc Teamwork: \\
Ambiguous Tasks, Roles, and Communication}

\numberofauthors{4}

\author{
\alignauthor
Jonathan Grizou\\
      \affaddr{Flowers Team}\\
      \affaddr{INRIA - ENSTA ParisTech}\\
      \affaddr{France}\\
      \email{jonathan.grizou@inria.fr}
\alignauthor
Samuel Barrett\\
      \affaddr{Kiva Systems}\\
      \affaddr{North Reading, MA 01864 USA}\\
      \email{basamuel@kivasystems.com}
\and
\alignauthor
Peter Stone\\
      \affaddr{Dept. of Computer Science}\\
      \affaddr{The Univ. of Texas at Austin}\\
      \affaddr{Austin, TX 78712 USA}\\
      \email{pstone@cs.utexas.edu}
\alignauthor
Manuel Lopes\\
      \affaddr{Flowers Team}\\
      \affaddr{INRIA - ENSTA ParisTech}\\
      \affaddr{France}\\
      \email{manuel.lopes@inria.fr}
}

\newcommand{\env}{E}
\newcommand{\task}{\tau}
\newcommand{\config}{\kappa}
\newcommand{\com}{\rho}
\newcommand{\gridwidth}{1}
\newcommand{\figwidth}{1}

@@

\maketitle
\begin{abstract}
% Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents.
Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. We consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that our ad hoc agent can join such a team without barely impacting the overall performance compared to a pre-coordinated agent.
\end{abstract}

@@

\bibliographystyle{abbrv}
\bibliography{adhoccom}

### 2016-01-31 paper/domain.tex
formatting for http://ala2016.csc.liv.ac.uk/

Each position on the grid is called a state $s$, which for convenience is also described as the $(x,y)$ coordinate. For each domain hypothesis $d \in D$, the environment $\env$ is the same, including its dynamic and noise level. A task $\task$ is fully defined by the position of the capture state, denoted $s_C$, that could be any grid cell. The reward function is one when the prey is locked on that state and is zero otherwise. A team configuration $\conf$ describes the role of each agent. For example, $\conf = [N,E,S,W]$ indicates that the first agent is in charge of the North side of the prey, the second one of the East side, etc. The communication protocol $\com$ includes a mapping and a reference (more details are provided in Section~\ref{sec:com}).

>>>>

Each position on the grid is called a state $s$, which for convenience is also described as the $(x,y)$ coordinate. For each domain hypothesis $d \in D$, the environment $\env$ is the same, including its dynamic and noise level. A task $\task$ is fully defined by the position of the capture state, denoted $s_C$, that could be any grid cell. The reward function is one when the prey is locked on that state and is zero otherwise. A team configuration $\config$ describes the role of each agent. For example, $\config = [N,E,S,W]$ indicates that the first agent is in charge of the North side of the prey, the second one of the East side, etc. The communication protocol $\com$ includes a mapping and a reference (more details are provided in Section~\ref{sec:com}).

### 2016-01-31 paper/problem.tex
formatting for http://ala2016.csc.liv.ac.uk/

\begin{itemize}
\item An \textbf{environment} $\env$ made of $n_S$ states, which we denote $\{s_1, \ldots, s_{n_S}\}$, and where agents can perform $n_U$ actions, which we denote $\{u_1, \ldots, u_{n_U}\}$. The environment dynamics are known and described by a probability distribution that for any given state $s$ and action $u$ gives the probability of a next state $s'$, $p(s'|s,u,\env)$.
\item A \textbf{task} $\task$ that the agents should achieve, represented by a reward function $R$.
\item A \textbf{configuration} $\conf$ that defines the role given to each agent, i.e. their specialties.
\item A \textbf{protocol} $\com$ that defines the way agents communicate to each others, i.e. their language. We denote $m_{b}$ as the message of an agent $b$.
\end{itemize}

A domain is defined by $d = \{\env, \task, \conf, \com\}$ that is a subset of all possible domains $D$. We denote $S$ as the set of all agent states, $ S~=~\{ s_{b_1}, \ldots, s_{b_{n_B}} \} $, respectively $S'$ the set of all agent next states. We denote $M$ as the set of all agents' messages, $M~=~\{m_{b_1}, \ldots, m_{b_{n_B}}\}$. We want to evaluate how an ad hoc agent $a$ can adapt in such a domain. To evaluate its performance, we remove one agent randomly from a fully formed team, creating the set $B^-$, and replace it by the ad hoc agent. The resulting team is denoted as $B^{-}_{a}$.
%
The team performance is evaluated on the task $\task$ using the reward function $R$. We denote $score(B, d)$ as the score resulting from the team $B$ executing the problem $d$, i.e. the accumulated reward. In this work, we want to create an ad hoc agent that minimizes the score loss between the original team $score(B, d)$ and the team with the ad hoc agent $score(B^{-}_{a}, d)$. The problem is that the ad hoc agent needs to fit into a team yet unknown to it. It must therefore identify all the components of the domain ($\env, \task, \conf, \com$). The main challenge is that the ad hoc agent does not have direct access to its performance. Indeed, it cannot compute $score(B^{-}_{a}, d)$ because $d$ is unknown to it.

>>>>

\begin{itemize}
\item An \textbf{environment} $\env$ made of $n_S$ states, which we denote $\{s_1, \ldots, s_{n_S}\}$, and where agents can perform $n_U$ actions, which we denote $\{u_1, \ldots, u_{n_U}\}$. The environment dynamics are known and described by a probability distribution that for any given state $s$ and action $u$ gives the probability of a next state $s'$, $p(s'|s,u,\env)$.
\item A \textbf{task} $\task$ that the agents should achieve, represented by a reward function $R$.
\item A \textbf{configuration} $\config$ that defines the role given to each agent, i.e. their specialties.
\item A \textbf{protocol} $\com$ that defines the way agents communicate to each others, i.e. their language. We denote $m_{b}$ as the message of an agent $b$.
\end{itemize}

A domain is defined by $d = \{\env, \task, \config, \com\}$ that is a subset of all possible domains $D$. We denote $S$ as the set of all agent states, $ S~=~\{ s_{b_1}, \ldots, s_{b_{n_B}} \} $, respectively $S'$ the set of all agent next states. We denote $M$ as the set of all agents' messages, $M~=~\{m_{b_1}, \ldots, m_{b_{n_B}}\}$. We want to evaluate how an ad hoc agent $a$ can adapt in such a domain. To evaluate its performance, we remove one agent randomly from a fully formed team, creating the set $B^-$, and replace it by the ad hoc agent. The resulting team is denoted as $B^{-}_{a}$.
%
The team performance is evaluated on the task $\task$ using the reward function $R$. We denote $score(B, d)$ as the score resulting from the team $B$ executing the problem $d$, i.e. the accumulated reward. In this work, we want to create an ad hoc agent that minimizes the score loss between the original team $score(B, d)$ and the team with the ad hoc agent $score(B^{-}_{a}, d)$. The problem is that the ad hoc agent needs to fit into a team yet unknown to it. It must therefore identify all the components of the domain ($\env, \task, \config, \com$). The main challenge is that the ad hoc agent does not have direct access to its performance. Indeed, it cannot compute $score(B^{-}_{a}, d)$ because $d$ is unknown to it.

### 2016-01-31 paper/results.tex
formatting for http://ala2016.csc.liv.ac.uk/

\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.245\textwidth}
    \centering
    \captionsetup{width=0.95\textwidth}
    \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=\figwidth\textwidth]{plots/with_noise/teamComparaison.png}
    \caption{Comparison of teams with full observability (\emph{T-FO}), partial observability without (\emph{T-PO}) and with (\emph{T-POC}) communication. The use of communication in the partial observability case allows recovering similar performances than with full observability. %Only cases where no predators can see the prey reduce performances.
}
    \label{fig:cmpteam}
  \end{subfigure}
  \begin{subfigure}[t]{0.245\textwidth}
    \centering
    \captionsetup{width=0.9\textwidth}
    \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=\figwidth\textwidth]{plots/with_noise/fullObs.png}
    \caption{Comparison of default team (\emph{T-FO}), ad hoc team (\emph{A-FO}), or a team including a predator with random policy (\emph{R-FO}). All predators have full observability. The inclusion of our ad hoc agent does not impact the performance.}
    \label{fig:fullobs}
  \end{subfigure}
  \begin{subfigure}[t]{0.245\textwidth}
    \centering
    \captionsetup{width=1.15\textwidth}
    \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true,width=\figwidth\textwidth]{plots/with_noise/partialObsCom.png}
    \caption{Comparison of default team (\emph{T-POC}), ad hoc team (\emph{A-POC}), or a team with one muted (\emph{T-POC-OM}) or one non-communicanting (\emph{T-POC-ONC}) predator. All predators have partial observability. The ad hoc agent does not communicate. The inclusion of our ad hoc agent does not impact the performances compared to \emph{T-POC-OM}.}
    \label{fig:partialobscom}
  \end{subfigure}
  \begin{subfigure}[t]{0.245\textwidth}
    \centering
    \captionsetup{width=0.62\textwidth}
    \includegraphics[trim=1.5cm 0.3cm 2.7cm 1.8cm, clip=true,width=\figwidth\textwidth]{plots/with_noise/computationalTime.png}
    \caption{After 20 steps, most hypotheses are discarded and the updates become faster. \emph{A-FO} (respectively \emph{A-POC}) disambiguates between 100 (respectively 1000) hypotheses.}
    \label{fig:comptime}
  \end{subfigure}
  \caption{(a) (b) (c) Average number of preys captured through steps for different conditions. (d) Computational time per step (log scale) for the ad hoc agent. All results include 20 percent of noise in actions and communications.\vspace{-0.3cm}}
  \label{fig:results}
\end{figure*}

@@

\vspace{-0.3cm}
\paragraph{Default Team Performance}

@@

@@

\vspace{-0.3cm}
\paragraph{Ad Hoc with Full Observability}

@@

@@

\vspace{-0.3cm}
\paragraph{Ad Hoc with Partial Observability}

@@

@@

>>>>

% \begin{figure*}[ht]
%   \centering
%   \begin{subfigure}[t]{0.245\textwidth}
%     \centering
%     \captionsetup{width=0.95\textwidth}
%     \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=\figwidth\textwidth]{plots/with_noise/teamComparaison.png}
%     \caption{Comparison of teams with full observability (\emph{T-FO}), partial observability without (\emph{T-PO}) and with (\emph{T-POC}) communication. The use of communication in the partial observability case allows recovering similar performances than with full observability. %Only cases where no predators can see the prey reduce performances.
% }
%     \label{fig:cmpteam}
%   \end{subfigure}
%   \begin{subfigure}[t]{0.245\textwidth}
%     \centering
%     \captionsetup{width=0.9\textwidth}
%     \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=\figwidth\textwidth]{plots/with_noise/fullObs.png}
%     \caption{Comparison of default team (\emph{T-FO}), ad hoc team (\emph{A-FO}), or a team including a predator with random policy (\emph{R-FO}). All predators have full observability. The inclusion of our ad hoc agent does not impact the performance.}
%     \label{fig:fullobs}
%   \end{subfigure}
%   \begin{subfigure}[t]{0.245\textwidth}
%     \centering
%     \captionsetup{width=1.15\textwidth}
%     \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true,width=\figwidth\textwidth]{plots/with_noise/partialObsCom.png}
%     \caption{Comparison of default team (\emph{T-POC}), ad hoc team (\emph{A-POC}), or a team with one muted (\emph{T-POC-OM}) or one non-communicanting (\emph{T-POC-ONC}) predator. All predators have partial observability. The ad hoc agent does not communicate. The inclusion of our ad hoc agent does not impact the performances compared to \emph{T-POC-OM}.}
%     \label{fig:partialobscom}
%   \end{subfigure}
%   \begin{subfigure}[t]{0.245\textwidth}
%     \centering
%     \captionsetup{width=0.62\textwidth}
%     \includegraphics[trim=1.5cm 0.3cm 2.7cm 1.8cm, clip=true,width=\figwidth\textwidth]{plots/with_noise/computationalTime.png}
%     \caption{After 20 steps, most hypotheses are discarded and the updates become faster. \emph{A-FO} (respectively \emph{A-POC}) disambiguates between 100 (respectively 1000) hypotheses.}
%     \label{fig:comptime}
%   \end{subfigure}
%   \caption{(a) (b) (c) Average number of preys captured through steps for different conditions. (d) Computational time per step (log scale) for the ad hoc agent. All results include 20 percent of noise in actions and communications.\vspace{-0.3cm}}
%   \label{fig:results}
% \end{figure*}

@@

% \vspace{-0.3cm}
\paragraph{Default Team Performance}

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/teamComparaison.png}
  \caption{Comparison of teams with full observability (\emph{T-FO}), partial observability without (\emph{T-PO}) and with (\emph{T-POC}) communication. The use of communication in the partial observability case allows recovering similar performances than with full observability.}
  \label{fig:cmpteam}
\end{figure}

@@

% \vspace{-0.3cm}
\paragraph{Ad Hoc with Full Observability}

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/fullObs.png}
  \caption{Comparison of default team (\emph{T-FO}), ad hoc team (\emph{A-FO}), or a team including a predator with random policy (\emph{R-FO}). All predators have full observability. The inclusion of our ad hoc agent does not impact the performance.}
  \label{fig:fullobs}
\end{figure}

@@

% \vspace{-0.3cm}
\paragraph{Ad Hoc with Partial Observability}

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/partialObsCom.png}
  \caption{Comparison of default team (\emph{T-POC}), ad hoc team (\emph{A-POC}), or a team with one muted (\emph{T-POC-OM}) or one non-communicanting (\emph{T-POC-ONC}) predator. All predators have partial observability. The ad hoc agent does not communicate. The inclusion of our ad hoc agent does not impact the performances compared to \emph{T-POC-OM}.}
  \label{fig:partialobscom}
\end{figure}

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/computationalTime.png}
  \caption{After 20 steps, most hypotheses are discarded and the updates become faster. \emph{A-FO} (respectively \emph{A-POC}) disambiguates between 100 (respectively 1000) hypotheses.}
  \label{fig:comptime}
\end{figure}
